Geometric variational problems

### 2020-11-13 NSF2020/Description/preliminarywork.tex
Update preliminarywork.tex
Preliminary section done; please check how table looks.

The PIs BQ, RR, YNY and Peter Fu initiated a collaboration in summer, 2020. 
The present proposal is the outgrowth of this collaboration. We have created
a compact numerical solver and included background flow in the simulations. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%For Specific Aim 3
%The background flow enters by replacing the third equation in \eqref{eq:stokes} 
%with the condition $(\mathbf{u} -\mathbf{u}_{\infty})(\mathbf{x}) 
%\to \mathbf{0}$ as $|\mathbf{x}| \to \infty$. To incorporate the far-field flow, 
%and using the representation 
%\begin{align}
%\label{PowerMiranda}
%  {\bf u} = {\bf u}_{\infty} + K\boldsymbol{\eta} + 
%    \sum_{i=1}^N S(\cdot, {\bf a}_i) {\bf F}_i + 
%                 R(\cdot, {\bf a}_i) {\bf G}_i.
%\end{align}
%The symbols $S$ and $R$ are Stokeslets and rotlets supported at the respective
%particle centers \cite{leal_2007} and $K\boldsymbol{eta}$ is a layer potential for the
%unknown density function $\boldsymbol{\eta}.$ The
%representation~\eqref{PowerMiranda} automatically satisfies all
%equations, with the exception of the rigid motion conditions. The rigid
%motion conditions follow by requiring the viscous stress vanishes across
%the particle boundaries.

@@

\subsubsection{Two-dimensional vesicles in shear flow}
We have simulated two-dimensional vesicles in the background shear flow
$\mathbf{u}_{\infty} = \dot{\gamma} x_2 \mathbf{i}_1$. The numerical solver
is written in the Matlab language runs on a desktop without parallelization. 
The initially circular vesicle bilayers rotate and elongate with the shear flow, 
and approach a steadily tank-treading ellipse.

There are a number of well-established mathematical models of vesicles
in shear flow \cite{torres-sanchez_millan_arroyo_2019, mahapatra_saintillan_rangamani_2020, 
Steigmann99, C6SM02452A}. We are trying to assess which constituative
assumption the particle-based bilayer satisfies. 
Our preliminary results show that the particle-based bilayers have the same
large area modulus as real lipid bilayer membranes, so we expect
realistic results from our flow simulations. Under moderate shear rates,
the vesicle elongates and takes on the tank-treading elliptical shape.
We can directly check for area compressibility conservation properties
from the particle simulation by tracking the change in area density of
the midplane as a function of shear flow rate, along with evaluating the
tangential divergence of the velocity.

Bilayer membranes have a small permeability to water
\cite{323e9a2f0c58487ea82518d7a1f96485}, and modelers often assume a
vesicle membrane that conserves volume. In our particle-based approach,
water motion across the bilayer is limited by the size of the gaps
between the particles, which is an artifact of coarse-graining.  Making
these gaps smaller comes at the expense of numerical accuracy, and we
will assess if there is a reasonable trade-off between approximate
volume conservation and simulation complexity.

Finally, the particle-based vesicle bilayers have two distinct leaflets
consisting of the particles in contact with the vesicle interior, and
those in contact with the surrounding fluid. Since there is a nonzero
separation between these layers, an important effect we observe is that
the leaflets slide against one another under shear flow.  This implies
that part of the viscous dissipation in the aqueous phase is enhanced by
intermonolayer friction~\cite{SHKULIPA2005823, ShkulipaThesis}.
Intermonolayer slip enters zero-thickness membrane models as a velocity
jump boundary condition and leads to multiple solution
states~\cite{schwalbe_vlahovska_miksis_2010}.  In the present proposal,
intermonolayer slip is an artifact of monolayer independence  and we can
control for slip as a function of shear rate, vesicle diameter, and
particle geometry.

>>>>

The PIs BQ, RR and YNY initiated a collaboration in summer, 2020 to develop
the hydrodynamic model of \S \ref{subsec:specific_aim_3}
To date, we have written solvers for the two-dimensional boundary integral formulatons
of \eqref{SL} and \eqref{eq:stokes}. The solvers gives third-order spatial accuracy
up to the boundary for $u$ and $\mathbf{u}$. The specialized reciprocal identity \ref{eq:reciprocal}
gives spectral accuracy for force and torque (Table \ref{tab:spectral_force}}).
\begin{wraptable}[10]{l}{0.45\textwidth}
\begin{center}
  \begin{tabular}{|l|l|l|l|}
 $n$   & $\max_{i} \|\mathbf{F}_i-\mathbf{F}_i^{\text{exact}}\|$   &  $\max_{i} \|\mathbf{G}_i-\mathbf{G}_i^{\text{exact}}\|$   \\
 32   &  81.4933     &   70.3970\\       
  64  &   4.9328     &    3.1928 \\
 128  &   0.0125     &    0.0030 \\
256   &   0.0000765   &   0.00003389        
  \end{tabular}
\end{center}
\caption{\label{tab:spectral_force} 
\footnotesize 
Numerical error for force and torque as a function of number of grid points $n$ per particle.
The data is for $N = 5$ particles characteristic diamater 2; two of the particles are nearly touching 
at a distance 0.02.} 
\end{wraptable}

We have simulated two-dimensional vesicles in shear flow 
$\mathbf{u}_{\infty} = \dot{\gamma} x_2 \mathbf{i}_1$ with shear rate $\dot{gamma}$. 
The particle-based system mimiks the behavior observed in continuous vesicles
\cite{torres-sanchez_millan_arroyo_2019, mahapatra_saintillan_rangamani_2020, 
Steigmann99, C6SM02452A}. The basic pattern is that at moderate shear rates,
the initially circular vesicle bilayer elongates and approaches a steadily 
tank-treading ellipse. Analyzing any underlying constituative laws, the particles
reorganize so that there is an initial decrease in vesicle area that plateaus 
to a steady state area, respectively increase in arc length. This suggests that
the particle system behaves as a semipermeable vesicle \cite{323e9a2f0c58487ea82518d7a1f96485,YAO2017728}.
Our preliminary results already showed that the particle-based bilayers have the same
large area modulus as real lipid bilayer membranes, and permit a small amount of stretching. 
For large shear rates, the vesicle ruptures and flow carries away segmented membrane patches.

The above considerations justify using the particle-based model to study material properties 
as outline in \S \ref{subsec:specific_aim_1}. Furthermore, the particle system allows
for intermonolayer slip~\cite{SHKULIPA2005823, ShkulipaThesis}.
Intermonolayer slip enters zero-thickness membrane models as a velocity
jump boundary condition \cite{schwalbe_vlahovska_miksis_2010}, for example, 
but in the present model is a consequence of monolayer independence.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%For Specific Aim 3
%The background flow enters by replacing the third equation in \eqref{eq:stokes} 
%with the condition $(\mathbf{u} -\mathbf{u}_{\infty})(\mathbf{x}) 
%\to \mathbf{0}$ as $|\mathbf{x}| \to \infty$. To incorporate the far-field flow, 
%and using the representation 
%\begin{align}
%\label{PowerMiranda}
%  {\bf u} = {\bf u}_{\infty} + K\boldsymbol{\eta} + 
%    \sum_{i=1}^N S(\cdot, {\bf a}_i) {\bf F}_i + 
%                 R(\cdot, {\bf a}_i) {\bf G}_i.
%\end{align}
%The symbols $S$ and $R$ are Stokeslets and rotlets supported at the respective
%particle centers \cite{leal_2007} and $K\boldsymbol{eta}$ is a layer potential for the
%unknown density function $\boldsymbol{\eta}.$ The
%representation~\eqref{PowerMiranda} automatically satisfies all
%equations, with the exception of the rigid motion conditions. The rigid
%motion conditions follow by requiring the viscous stress vanishes across
%the particle boundaries.

@@
